Reliable Inferences from Simply Parameterised Models

### 2012-08-29 ms.tex
Priors

\subsection{The Simply-Parameterised Model}
The model for the shape of the line is:
\begin{eqnarray}
\mathcal{M}(x; A, c, w) &=&
A\exp
\left[
-\frac{1}{2w^2}
\left(x - c\right)^2
\right]
\end{eqnarray}
where $A$ is the peak amplitude, $c$ is the central position of the line
and $w$ is the line width. To infer the parameters $\{A, c, w\}$ from some data
$\yy$, we will use Bayes' rule:
\begin{eqnarray}
p(A, c, w | \yy) \propto p(A, c, w)p(\yy | A, c, w).
\end{eqnarray}
To proceed, we just need to define the prior and the sampling distribution.

@@

\begin{eqnarray}
p(\yy|A, c, w) &=& \prod_{i=1}^N
\frac{1}{\sigma\sqrt{2\pi}}
\exp
\left[
-\frac{1}{2\sigma^2}\left(y_i - \mathcal{M}(x_i; A, c, w)\right)^2
\right] \\
&\propto& \exp\left[-\frac{1}{2\sigma^2}\sum_{i=1}^N\left(y_i - \mathcal{M}(x_i; A, c, w)\right)^2\right] \\
&\propto& \exp\left[-\frac{1}{2}\chi^2\right].
\end{eqnarray}

>>>>

\section{Fitting a Simply-Parameterised Model}
We now explore what happens when we fit simulated data (bottom panel of
Figure~\ref{emission_line}) with an inflexible simply-parameterised model.
We will fit the data with single Gaussian component.
The model for the shape of the line is:
\begin{eqnarray}
\mathcal{M}(x; A, c, w) &=&
A\exp
\left[
-\frac{1}{2w^2}
\left(x - c\right)^2
\right]
\end{eqnarray}
where $A$ is the peak amplitude, $c$ is the central position of the line
and $w$ is the line width. To infer the parameters $\{A, c, w\}$ from some data
$\yy = \{y_1, y_2, ..., y_N\}$, we will use Bayes' rule:
\begin{eqnarray}
p(A, c, w | \yy) \propto p(A, c, w)p(\yy | A, c, w).
\end{eqnarray}
To proceed, we just need to define the prior and the sampling distribution.

@@

\begin{eqnarray}
p(\yy|A, c, w) &=& \prod_{i=1}^N
\frac{1}{\sigma\sqrt{2\pi}}
\exp
\left[
-\frac{1}{2\sigma^2}\left(y_i - \mathcal{M}(x_i; A, c, w)\right)^2
\right] \\
&\propto& \exp\left[-\frac{1}{2\sigma^2}\sum_{i=1}^N\left(y_i - \mathcal{M}(x_i; A, c, w)\right)^2\right] \\
&\propto& \exp\left[-\frac{1}{2}\chi^2({\tt model}, {\tt data})\right].
\end{eqnarray}
The likelihood here is related to the commonly used $\chi^2$ deviation between
te model and the data.

\subsection{The Prior Distribution}
To proceed with the inference, we require a prior distribution over the space
of possible parameter values for $A$, $c$ and $w$. We shall assume independence:
\begin{eqnarray}
p(A, c, w) = p(A)p(c)p(w).
\end{eqnarray}
This means that, prior to learning the data, we are unaware of any logical
connection between the values of the parameters. In other words, if we learned the
value of any parameter we would not update our knowledge about the others.
For the central position $c$ of the line, we assigned a uniform prior between
$-10$ and $10$, which is the same range as the data. For $A$ and $w$ we assigned
priors proportional to $1/A$ and $1/w$ between $10^{-3}$ and $10^{+3}$, since these
parameters are known to be positive and have an uncertain magnitude.
